Efficient FPGA Implementation of\\ Digit Parallel Online Arithmetic Operators

### 2014-06-03 FPT2014/bare_conf.tex
Draft 2nd version

\begin{abstract}
%\boldmath
Online arithmetic has been widely studied for ASIC implementation. Online components are designed to perform computations most significant digit first, resulting in the ability to chain arithmetic operators together for low latency. Recent research also shows that digit parallel online operators can fail more gracefully when operating beyond the deterministic clocking region, in comparison to operators with conventional arithmetic. Unfortunately, the utilization of online arithmetic operators in the past has required a large area overhead for FPGA implementation. In this paper, we propose novel approaches to implement the key primitives of online arithmetic, adders and multipliers, efficiently on modern Xilinx FPGAs with 6-input LUTs and carry resources. We demonstrate experimentally that in comparison to a direct implementation, the proposed architectures achieve over $67\%$ slice savings and over $1.2\times$ speed-ups for adders, and more than $3\times$ area savings and $1.5\times$ speed-ups for multipliers. Furthermore in comparison to standard arithmetic primitives, the area overheads of online operators drop from $8.41\times$ to $1.88\times$ for adders, and $8.11\times$ to $0.54\times$ for multipliers using our approach.

@@

In this paper, both issues will be addressed. In Section~\ref{Sec:OA_FPGA}, we propose a novel approach to map the digit parallel online adders onto FPGAs with 6-LUTs. The mapping is based on the fast carry logic, and the available resources within a slice can be fully utilized. In Section~\ref{Sec:OM_FPGA}, we optimize the online multiplication algorithm to yield an efficient FPGA implementation of a digit parallel online multiplier. The resource usage of both operators can be determined analytically. We demonstrate experimentally on Xilinx Virtex-6 FPGAs that the proposed designs achieve significant area reduction and performance gain over their original implementations. For online adders, our method achieves over $67\%$ slice savings and $1.2\times$ frequency speed-ups in comparison to a direct implementation. The area overhead drops from up to $8.41\times$ to $1.88\times$ using the proposed online adder in comparison to the standard ripple-carry adder. For online multipliers, the LUTs and slice utilization of the proposed new architecture drops over $2.4\times$ and $3.2\times$, with a frequency speed-up more than $1.5\times$ in comparison to a direct implementation. When compared to the multiplier with standard arithmetic, the direct implementation has an area overhead up to $8.11\times$. This value drops to $1.84\times$ when using the proposed online multiplier. Furthermore, using the proposed structure can even achieve area savings of up to $55\%$ for half precision results in comparison to the standard multiplier.

>>>>

\begin{abstract}
%\boldmath
Online arithmetic has been widely studied for ASIC implementation. Online components are designed to perform computations most significant digit first, resulting in the ability to chain arithmetic operators together for low latency. Recent research also shows that digit parallel online operators can fail more gracefully when operating beyond the deterministic clocking region, in comparison to operators with conventional arithmetic. Unfortunately, the utilization of online arithmetic operators in the past has required a large area overhead for FPGA implementation. In this paper, we propose novel approaches to implement the key primitives of online arithmetic, adders and multipliers, efficiently on modern Xilinx FPGAs with 6-input LUTs and carry resources. We demonstrate experimentally that in comparison to a direct implementation, the proposed architectures achieve over $67\%$ slice savings and over $1.2\times$ speed-ups for adders, and more than $3\times$ slice savings and $1.5\times$ speed-ups for multipliers. Furthermore in comparison to standard arithmetic primitives, the area overheads of online operators drop from $8.41\times$ to $1.88\times$ for adders, and $8.11\times$ to $0.54\times$ for multipliers using our approach.

@@

In this paper, both issues will be addressed. In Section~\ref{Sec:OA_FPGA}, we propose a novel approach to map the digit parallel online adders onto FPGAs with 6-LUTs. The mapping is based on the fast carry logic, and the available resources within a slice can be fully utilized. In Section~\ref{Sec:OM_FPGA}, we optimize the online multiplication algorithm to yield an efficient FPGA implementation of a digit parallel online multiplier. The resource usage of both operators can be determined analytically. We demonstrate experimentally on Xilinx Virtex-6 FPGAs that the proposed designs achieve significant area reduction and performance gain over their original implementations. For online adders, our method achieves over $67\%$ slice savings and $1.2\times$ frequency speed-ups in comparison to a direct implementation. The area overhead drops from up to $8.41\times$ to $1.88\times$ using the proposed online adder in comparison to the standard ripple-carry adder. For online multipliers, the LUTs and slice utilization of the proposed new architecture drops over $2.4\times$ and $3.2\times$, with a frequency speed-up more than $1.5\times$ in comparison to a direct implementation. When compared to the multiplier with standard arithmetic, the direct implementation has an area overhead up to $8.11\times$. This value drops to $1.84\times$ when using the proposed online multiplier. In addition, using the proposed structure can even achieve area savings of up to $55\%$ for half precision results in comparison to the standard multiplier.
